---
title: "Homework 6"
output: html_notebook
---

`Prepare a 3-4 page report describing the probability of survival on the Titanic as the function of age, sex, class, and the price the passengers paid for the ticket.`

`The report should contain the description of the data using tables/graphs, the outline of the hypotheses, and their results, followed by the conclusion.`

`Your report should be formatted and presented as if it were a paper, not just a collection of random facts.`

`Submit both the Rmd and the HTML file.`

```{r}
library(tidyverse)
```

```{r}
titanic <- read_csv("titanic.csv")

titanic1 <- titanic %>% mutate(surv = ifelse(Survived == 1, "Yes", "No"),
                               class = case_when(Pclass == 1 ~ "First",
                                                 Pclass == 2 ~ "Second",
                                                 Pclass == 3 ~ "Third"))
```

H0: Reason of survival was unrelated with the age of the passengers, class, sex and price.

H1: Reason of survival is related with the age of passengers, class, sex, and price.

Now I want to test my null hypothesis:

**survival** as a function of **age, class, sex, price**

```{r}
titanic1 %>% ggplot(aes(x=Survived, y=class)) + geom_point()

plot(titanic)
plot(titanic1)
```

clean the titanic1 dataset to only the necessary functions (=survival, age, sex, fare, pass class).

```{r}
titanic2 <- titanic1 %>% select(Age, Sex, Fare, Pclass, Survived)

plot(titanic2)
```

```{r}
titanic3 <- lm(Survived~Age + Sex + Pclass + Fare, data = titanic2)
summary(titanic3)
```

```{r}

```
