---
title: "Homework 3"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
[SYB61_T03_Population Growth Rates in Urban areas and Capital cities.csv](https://github.com/LoizosKo/helicopter/files/7741996/SYB61_T03_Population.Growth.Rates.in.Urban.areas.and.Capital.cities.csv)
1. Examined the file in R. How many unique countries are there? What variables are available? Are the data in long form or in wide form?
2. You need to make sure that only one type of information is stored in each column. If that is not the case, transform the data using applicable functions (verbs) from the tidyr package. 
3. Now you want to merge the data with the gapminder dataset. Let's work with data from 2007 (so you might want to filter out the rest of the data). The problem is that your dataset with more refined population estimates does not have data for 2007, but it has data for 2005 and 2010. Can you somehow interpolate the data (i.e. predict what the values were in 2007)?  You might approach the problem in different way, but make sure to describe exactly what you did.
4. After you merged the data, you can actually check how well you predicted the population. Remember that the gapminder dataset already has the population data for 2007. Calculate the absolute and the relative error of prediction using your technique of interpolation for each country in the data. What is the smallest error? What is the largest error? How on average did your interpolation perform?
5. Calculate the proportion of the population in each country that live in the capital city. Give summary statistics for the distribution of this proportion. Plot the graph showing how gdp per capita is linked to the proportion of people living in a capital city. Plot a similar graph linking life expectancy and the proportion of people living in capital cities. Think about what is more logical to place on x and y axes.
6. Plot the histogram of proportion of people living in rural areas in 2007 and the histogram of the proportion of people living in capital cities. Talk about summary statistics of these variables
```{r}
library(tidyverse)
library(readxl)
library(gapminder)

```

```{r}
hwdata <- read_excel("hw3data.xlsx", 
                     col_names = FALSE, skip = 2)
```

```{r}
colnames(hwdata) <- c("number","country", 
                      "year", "series", "capital", "capital city footnote", 
                      "value", "footnotes", "source") 
```

1.  `Examined the file in R. How many unique countries are there? What variables are available? Are the data in long form or in wide form?`

**In the following chunks (line 53 to 78) we can see that there are in total 137 countries for this analysis. Data are in wide form because they have many columns and a few rows. In line 80 to 84 we can see that there are 9 variables.**

**Want to see how many areas there are**

```{r}
hwdata %>% distinct(`country`)
#Running this code shows us that there are 261 areas
```

**Named 'capital city' to 'capital'**

```{r}
capitaldata <- hwdata %>% distinct(capital)

capc <- hwdata%>% filter(capital!="")
```

**For the purpose of this to find the number of unique countries, I will assume that the unique countries are those described in gapminder dataset.**

```{r}
hwdata%>% distinct(country)
gapminder%>% distinct(country)
gapminder%>% filter(year==max(year)) %>%
  inner_join(hwdata, by="country")
```

**\#anti_join shows us the data that failed to match with 'gapminder' dataset**

```{r}
gapminder%>% filter(year==max(year)) %>%
  anti_join(hwdata, by = "country")

#useful for exercise 3 afterwards
gap2007 <- gapminder%>% filter(year %in% c(2007)) 
gap2007 %>% anti_join(totmerg2005, by = "country")
#Reunion, Taiwan, West Bank and Gaza, Swaziland, Taznania
```

**Here are the data did not match with the 'hwdata' dataset.**

```{r}
hwdata %>%
  anti_join(filter(gapminder, year==max(year)), by = "country") %>%
  distinct(country)
```

**Before merging 'gapminder' with 'hwdata' datasets, I make sure that they have identical country names so they can be matched properly.**

```{r}
mergedcountry <- gapminder %>% mutate(country= 
                       case_when(country== "United States" ~ "United States of America",
                                        country== "Venezuela" ~ "Venezuela (Boliv. Rep. of)",
                                        country== "Yemen, Rep" ~ "Yemen",
                                        country== "Vietnam" ~ "Viet Nam",
                                        country== "Syria" ~ "Syrian Arab Republic",
                                        country== "Slovak Republic" ~ "Slovakia",
                                        country== "Korea, Rep." ~ "Republic of Korea",
                                        country== "Korea, Dem. Rep." ~ "Dem. People's Rep. Korea",
                                        country== "Iran" ~ "Iran (Islamic Republic of)",
                                        country== "Hong Kong, China" ~ "China, Hong Kong SAR",
                                        country== "Czech Republic" ~ "Czechia",
                                        country== "Cote d'Ivoire" ~ "CÙte díIvoire",
                                        country== "Congo, Rep." ~ "Congo",
                                        country== "Congo, Dem. Rep." ~ "Dem. Rep. of the Congo",
                                        country== "Bolivia" ~ "Bolivia (Plurin. State of)",
                                        country== "Reunion" ~ "RÈunion",
                                        TRUE ~ as.character(country)))  %>%
  
  #Now that I matched the country names that I want to include, I use the following code to see which ones I matched in out of 261.
  filter(year==max(year)) %>%
  inner_join(hwdata, by = "country") %>%
  #anti_join(hwdata, by = "country") %>%, this command would show which countries were not matched with gapminder
  distinct(country)
```

**hwdata %\>% ls() shows us that there are 9 variables**

```{r}
hwdata %>% ls()
```

2.  `You need to make sure that only one type of information is stored in each column. If that is not the case, transform the data using applicable functions (verbs) from the tidyr package. (we will cover it on Monday)`

**In line 90 to 92, I use pivot_wider to flip the data in the 'series' column vertically and match them with the 'value' column**

```{r}
hwdata %>% pivot_wider(id_cols = c(country, year), names_from = series, values_from = value) %>%
  pivot_wider(id_cols = c())
```

3.  `Now you want to merge the data with the gapminder dataset. Let's work with data from 2007 (so you might want to filter out the rest of the data). The problem is that your dataset with more refined population estimates does not have data for 2007, but it has data for 2005 and 2010. Can you somehow interpolate the data (i.e. predict what the values were in 2007)? You might approach the problem in different way, but make sure to describe exactly what you did.`

`HINT: think about assumptions that need to be made about the changes in population from year to year. How likely is it that these changes will be very drastic? What can cause drastic changes? If you do not expect drastic causes, can the population just change slowly and steadily?`

**The data are cleaned at this chunk so we have values for 2005 and 2010.**

```{r}
hwdata %>%
  inner_join(filter(gapminder, year==max(year)), by = "country") %>%
  distinct(country)
gap2007 <- gapminder%>% filter(year %in% c(2007)) 

hwdata %>% filter(year %in% c(2005))
hwdata %>% filter(year %in% c(2010)) 

tot <- merge(mergedcountry,hwdata,by="country") %>%
  pivot_wider(id_cols = c(country, year),
              names_from = series,
              values_from = value) 

total <- merge(mergedcountry,hwdata,by="country") %>%
  filter(year %in% c(2005, 2010))

totmerg <- total %>% pivot_wider(id_cols = c(country, year),
              names_from = series,
              values_from = value) 

colnames(totmerg) <- c("country","year", 
                      "capcitpop", "capcitpopperurbpop", "urbpopper", "capcitpoppertotpop", 
                      "rurpoppergrorat", "urbpoppergrorat") 

```

**First step to calculate the 2007 population, is to create a population column in the totmerg dataset.**

```{r}
totmergpop <- totmerg %>% mutate(totpop = (capcitpop/capcitpoppertotpop*100), 
                                 totrurpop = totpop * (1-urbpopper/100),
                                 toturbpop = totpop * (urbpopper /100))
```

**Now we take the urban and rural growth rates between two ranging years (2005 and 2010) to find data for 2007. Then we sum up the numbers for urban and rural to find the total pop in 2007.**

**When we look at the data, in the column tot2007, we can see the total population in 2007 next to the rows of the countries in 2005. This will also help for doing the histogram in question 6.**

```{r}
totmertot<- totmergpop %>% mutate(totrur2007 = totrurpop * (1+rurpoppergrorat/100)^2*1000,
                      toturb2007 = toturbpop * (1+urbpoppergrorat/100)^2*1000,
                      tot2007 = (toturb2007 + totrur2007)) %>%  filter(year %in% c(2005))
```

**We begin by removing the unmatched country rows from the gapminder dataset (Taiwan, Swaziland, Taznania, West Bank and Gaza, and Yeamen, Rep.**

```{r}
gap2007match<- 
  subset(gap2007, country!="Taiwan"& 
         country!="Tanzania" & 
         country!="Swaziland" &
         country!="West Bank and Gaza"&
           country!="Yemen, Rep.") %>%
  mutate(country= 
                       case_when(country== "United States" ~ "United States of America",
                                        country== "Venezuela" ~ "Venezuela (Boliv. Rep. of)",
                                        country== "Yemen, Rep" ~ "Yemen",
                                        country== "Vietnam" ~ "Viet Nam",
                                        country== "Syria" ~ "Syrian Arab Republic",
                                        country== "Slovak Republic" ~ "Slovakia",
                                        country== "Korea, Rep." ~ "Republic of Korea",
                                        country== "Korea, Dem. Rep." ~ "Dem. People's Rep. Korea",
                                        country== "Iran" ~ "Iran (Islamic Republic of)",
                                        country== "Hong Kong, China" ~ "China, Hong Kong SAR",
                                        country== "Czech Republic" ~ "Czechia",
                                        country== "Cote d'Ivoire" ~ "CÙte díIvoire",
                                        country== "Congo, Rep." ~ "Congo",
                                        country== "Congo, Dem. Rep." ~ "Dem. Rep. of the Congo",
                                        country== "Bolivia" ~ "Bolivia (Plurin. State of)",
                                        country== "Reunion" ~ "RÈunion",
                                        TRUE ~ as.character(country)))

totmertot %>% mutate(country= 
                       case_when(country== "United States" ~ "United States of America",
                                        country== "Venezuela" ~ "Venezuela (Boliv. Rep. of)",
                                        country== "Yemen, Rep" ~ "Yemen",
                                        country== "Vietnam" ~ "Viet Nam",
                                        country== "Syria" ~ "Syrian Arab Republic",
                                        country== "Slovak Republic" ~ "Slovakia",
                                        country== "Korea, Rep." ~ "Republic of Korea",
                                        country== "Korea, Dem. Rep." ~ "Dem. People's Rep. Korea",
                                        country== "Iran" ~ "Iran (Islamic Republic of)",
                                        country== "Hong Kong, China" ~ "China, Hong Kong SAR",
                                        country== "Czech Republic" ~ "Czechia",
                                        country== "Cote d'Ivoire" ~ "CÙte díIvoire",
                                        country== "Congo, Rep." ~ "Congo",
                                        country== "Congo, Dem. Rep." ~ "Dem. Rep. of the Congo",
                                        country== "Bolivia" ~ "Bolivia (Plurin. State of)",
                                        country== "Reunion" ~ "RÈunion",
                                        TRUE ~ as.character(country)))
  

```

4.  `After you merged the data, you can actually check how well you predicted the population. Remember that the gapminder dataset already has the population data for 2007. Calculate the absolute and the relative error of prediction using your technique of interpolation for each country in the data. What is the smallest error? What is the largest error? How on average did your interpolation perform?`

    **First we insert the gapminder total population in the dataset**

    ```{r}
    totmertotal <- totmertot %>%mutate(gap2007pop = gap2007match$pop)
    ```

    ***Then, we calculate the absolute error [=gap2007pop-tot2007] and the relative error [=gap2007pop-tot2007)/gap2007pop]***

    ```{r}
    totmertotal <-totmertotal %>% mutate (abserr = totmertotal$tot2007 - totmertotal$gap2007pop)

    totmertotal <- totmertotal %>% mutate( rererr= ((totmertotal$tot2007 - totmertotal$gap2007pop)/totmertotal$gap2007pop))
    ```

    **As we can see, the smallest absolute is Czechia with 1.87 and the largest is India with 9.83**

    **smallest absolute error is Egypt with 1.433 and largest is Chechia with 1.83**

    ```{r}
    totmertotalllll <- totmertotal%>% mutate(abserr=abs(abserr)) %>%
      summarise(min=(abserr))

    totmertotalll <- totmertotal%>% mutate(rererr=abs(rererr)) %>%
      summarise(min=(rererr))
    ```

5.  `Calculate the proportion of the population in each country that live in the capital city. Give summary statistics for the distribution of this proportion.`

    **We assume that the percentage increase of people living in capitals from 2005 to 2010 happened gradually. Then we pivot table so we have the years in columns. Following to that we create the column with 2007 percentage through the weighted average formula.**

    ```{r}

    totmerg22222 <- totmerg %>% pivot_wider(id_cols = c(country),
                                            names_from = year,
                                            values_from = c(capcitpoppertotpop))

    totmerg22222 <- totmerg22222 %>% 
      mutate(pop_2007 = ((totmerg22222$`2005` *(3/5)) + (totmerg22222$`2010`*(2/5))))

    ```

    `Plot the graph showing how gdp per capita is linked to the proportion of people living in a capital city.`

    **We first insert to the totmertotal data the population per capita proportion.**

    ```{r}
    totmertotal <- totmertotal %>% mutate(popratpercap2007= totmerg22222$pop_2007) 

    ```

    **Then we multiply the population per capita proportion with the total population to find the amount of people that lived in capital cities in 2007. As we can see, the capital city population proportion did not change significantly due to the assumption that its growth was gradual. For instance, Afghanistan population percentage living in capital in 2007 was 11.52, which is between 2010's percent (11.6) and 2005's percent (11.4)**

    ```{r}
    totmertotal<- totmertotal %>% mutate(capcitypop= popratpercap2007/100*tot2007)

    totmertotal<- totmertotal%>% mutate(gdppercap = gap2007match$gdpPercap)
    ```

    **Following to that we plot the graph**

    ```{r}
    totmertotal %>%  ggplot(aes(x = gdppercap, y= popratpercap2007)) + geom_point() + geom_smooth(method="lm", col="green", lwd=0.5, se=F, orientation =NA, na.rm=FALSE, inherit.aes = TRUE) + theme_classic()
    ```

    `Plot a similar graph linking life expectancy and the proportion of people living in capital cities. Think about what is more logical to place on x and y axes.`

    ```{r}
    totmertotal <- totmertotal%>%
      mutate(lifeexp= gap2007match$lifeExp)

    totmertotal%>%ggplot(aes(x = lifeexp, y= popratpercap2007)) + geom_point() + geom_smooth(method="lm", col="blue", lwd=0.5, se=F, orientation =NA, na.rm=FALSE, inherit.aes = TRUE) + theme_classic()
    ```

6.  `Plot the histogram of proportion of people living in rural areas in 2007 and the histogram of the proportion of people living in capital cities. Talk about summary statistics of these variables.`

**The amount of people living in rural and urban areas is known from exercise 3 -- it can be seen on the 'totmertotal'. This code helps us find the proportion.**

```{r}
totmertotal <- totmertotal%>% 
  mutate(rur2007per = totrur2007*100/tot2007,
         urb2007per = toturb2007*100/tot2007)
```

**This code does the histograms for the proportion of people living in rural areas and people living in capital cities.**

```{r}
#rural areas
totmertotal %>%
  ggplot(aes(x=rur2007per)) + geom_histogram() + geom_rug() + geom_density(aes(x=0.10, y=2.5), colour="red", adjust=30, sd=1)

#capital cities
totmertotal %>%
  ggplot(aes(x=popratpercap2007)) + geom_histogram() + geom_rug() + geom_density(aes(x=10, y=4), colour="red", adjust=3, sd=1)

```

**The capital cities histogram seems to be right skewed. In other words, the percentage of people living in capital cities is on the lowest percentages -- most of the countries proportion is between 5% and 25%. When it comes to people living in rural areas, the histogram seems that is tend to be more like normal distribution. In other words, the proportion of people living in rural areas is equally distributed among the countries.**

```{r}
hist(gapminder$lifeExp[gapminder$year == 1952])

gapminder %>% ggplot(aes(x=year, y=lifeExp)) + geom_point()+ geom_smooth(method="lm") 

gapminder %>%
    group_by(continent, year) %>%
    summarise(lifeExp=median(lifeExp))

gapminder %>%
    group_by(country) %>% filter(year=="1982") 
```

```{r}
gapminder %>% ggplot(aes(x=lifeExp, y=year))+ geom_point()+ geom_smooth(method="lm") 
```
