---
title: "Homeworks"
---

More about this website.



```{r}
library("gapminder")
library(tidyverse)
```

```{r}
gapminder <- gapminder
```

1.  I hereby examine the organization of the data set and randomly sample one country.

```{r}
library(dplyr)
set.seed(5098765)
randomcountry <- sample_n(gapminder, 1)
randomcountry
```
filter(year==2007)

2.  For the country that was sampled (Denmark), I demonstrate the dynamics of its population, life expectancy, and GDP per capita from the earliest to the latest year available.

-   Life expectancy, GDP per capita, and the population of Denmark gradually increased through the years.

```{r}
Denmark <- gapminder %>%
filter(country == "Denmark")
 Denmark %>%  ggplot(aes(x = year, y = pop)) + geom_line()
 Denmark %>%  ggplot(aes(x=year, y=lifeExp)) + geom_line()
 Denmark %>%  ggplot(aes(x=year, y=gdpPercap)) + geom_line() 
```

3. Show how does the population, life expectancy, and GDP per capita compare with the averages for the continent and for the world as a whole in the earliest year of observation and the latest year of observation. (You can hereby find my calculations of the corresponding means for this)

```{r}

#Mean Europe 1952
gapminder %>%   
  filter(continent=="Europe", year == min(year)) %>%
  summarise(avg_pop=mean(pop), avg_lifeExp=mean(lifeExp), avg_gdpPCap= mean(gdpPercap))

#Mean World 1952
gapminder %>%
  filter(year == min(year)) %>%
  summarise(avg_pop=mean(pop), avg_lifeExp=mean(lifeExp), avg_gdpPCap= mean(gdpPercap))

#Mean Europe 2007
gapminder %>%
  filter(continent=="Europe", year == max(year)) %>%
  summarise(avg_pop=mean(pop), avg_lifeExp=mean(lifeExp), avg_gdpPCap= mean(gdpPercap))

#Mean World 2007
gapminder %>%
  filter(year == max(year)) %>%
  summarise(avg_pop=mean(pop), avg_lifeExp=mean(lifeExp), avg_gdpPCap= mean(gdpPercap))


#For purposes of answering question 9:
#Mean Denmark 1952
gapminder %>%
  filter(country=="Denmark", year == min(year)) %>%
  summarise(avg_pop=mean(pop), avg_lifeExp=mean(lifeExp), avg_gdpPCap= mean(gdpPercap))

#Mean Denmark 2007
gapminder %>%
  filter(country=="Denmark", year == max(year)) %>%
  summarise(avg_pop=mean(pop), avg_lifeExp=mean(lifeExp), avg_gdpPCap= mean(gdpPercap))
```

4.  Calculate the standard deviation for the continent and for the world for each indicator for the first and last year of observations. You can see here how many standard deviations away from the mean was your country in the first year and in the last year of observation. This indicates that Denmark gradually developed through the years.

```{r}
#For Denmark (World) population
zscoreWorld<-gapminder%>% 
  group_by(year) %>%
  mutate(avgpop=mean(pop), sdpop=sd(pop))%>% 
  filter (country == "Denmark") %>%
  mutate(zscore=(pop-avgpop)/sdpop) %>%
  filter(year==min(year))
zscoreWorld 

#For Denmark (Continent) population
zscoreContinent<-gapminder%>% 
  group_by(year) %>%
  filter (continent == "Europe") %>%
  mutate(avgpop=mean(pop),sdpop=sd(pop))%>% 
  filter (country == "Denmark") %>%
  mutate(zscore=(pop-avgpop)/sdpop) %>%
  filter(year==min(year))
zscoreContinent

#For Denmark (World) Life Expectancy
zscoreWorldLifeExp<-gapminder%>% 
  group_by(year) %>%
  mutate(avglifeExp=mean(lifeExp), sdlifeExp=sd(lifeExp))%>% 
  filter (country == "Denmark") %>%
  mutate(zscore=(lifeExp-avglifeExp)/sdlifeExp) %>%
  filter(year==min(year))
zscoreWorldLifeExp

#For Denmark (Continent) Life Expectancy
zscoreContinentLifeExp<-gapminder%>% 
  group_by(year) %>%
  filter (continent == "Europe") %>%
  mutate(avglifeExp=mean(lifeExp),sdlifeExp=sd(lifeExp))%>% 
  filter (country == "Denmark") %>%
  mutate(zscore=(lifeExp-avglifeExp)/sdlifeExp) %>%
  filter(year==min(year))
zscoreContinentLifeExp

#For Denmark (World) GDP per Capita
zscoreWorldAvGDPpCap<-gapminder%>% 
  group_by(year) %>%
  mutate(avggdpPercap=mean(gdpPercap), sdgdpPercap=sd(gdpPercap))%>% 
  filter (country == "Denmark") %>%
  mutate(zscore=(gdpPercap-avggdpPercap)/sdgdpPercap) %>%
  filter(year==min(year))
zscoreWorldAvGDPpCap

#For Denmark (Continent) GDP per Capita
zscoreContinentAvGDPpCap<-gapminder%>% 
  group_by(year) %>%
  filter (continent == "Europe") %>%
  mutate(avggdpPercap=mean(gdpPercap),sdgdpPercap=sd(gdpPercap))%>% 
  filter (country == "Denmark") %>%
  mutate(zscore=(gdpPercap-avggdpPercap)/sdgdpPercap) %>%
  filter(year==min(year))
zscoreContinentAvGDPpCap


```